Global Climate and Atmosphere Composition of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and Spitzer Phase Curve Observations

### 2018-01-02 ms.tex
intro

\begin{abstract}
	We present thermal phase curve measurements for the hot Jupiter WASP-103b observed with \emph{Hubble}/Wide Field Camera 3 and \emph{Spitzer}/IRAC. The dayside planet-to-star flux ranges from FIXME to FIXME at $1.1 - 4.5\,\mu$m.  The phase curves are symmetric and large amplitude, indicative of poor heat redistribution. We fit the phase curves with the SPIDERMAN package to determine the brightness distribution in the photosphere, and find the peak dayside temperature ranges from FIXME to FIXME depending on model.  We also generate phase-resolved spectra, which are consistent with blackbody models for the planet's emission at near-infrared wavelengths, but show strong emission features in the \Spitzer\ bands. We attribute these features to CO emission and a thermal inversion. The retrieved atmospheric composition is more metal-enriched than other hot Jupiters (FIXME). The carbon-to-oxygen ratio is consistent with solar composition. The absence of water features is due to BLAH. We compare the phase curves to general circulation models and find that atmospheric drag is required.  We also compare the spectra to brown dwarfs and young directly imaged companions and find that the water features are significantly diminished in amplitude. power of phase curve observations to elucidate the three-dimensional structure of exoplanet atmospheres.
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
Planets are round, rotating, and irradiated on one hemisphere at a time -- all of which contribute to rich spatial structure in their climate and atmospheric composition.  Short period, tidally locked planets are an extreme example, with hot, continuously illuminated daysides. This asymmetry is expected to produce large gradients in temperature, chemistry, and cloud coverage with longitude \citep{showman09, kataria16, parmentier16}.

Extrasolar planets are so distant that they are not spatially resolved, so the only way to reveal this variation is to measure a phase curve. Phase curve observations consist of continuous monitoring of the planet-to-star flux ratio over a complete orbital revolution of the planet, which exposes different longitudes at each orbital phase.  The first phase curve was observed with \Spitzer\ for the hot Jupiter HD 189733b by \citep{knutson07}, followed by additional \Spitzer\ observations for about a dozen more systems \citep{parmentier17}.  \Kepler\ enabled the first measurements of phase curves at optical wavelengths \citep[e.g.]{esteves15}. \cite{stevenson14} observed the first \emph{spectroscopic} phase curve for WASP-43b with \emph{Hubble}/Wide Field Camera 3 (\HST/WFC3) in the near-infrared. These observations had fairly good agreement with 3D GCM predictions, eastward-shifted peak brightness and more efficient heat redistribution for cooler planets. 
FIXME: say after Spitzer that results agree with GCM. Elaborate after esteves15 that the optical phase curves are sensitive to clouds.

In this paper we present spectroscopic phase curve observations of the hot Jupiter WASP-103b. This planet is an ideal target for thermal phase curve observations, with an orbital period of 22 hours and an equilibrium temperature of 2500 K \citep{gillon14}. The dayside thermal emission of WASP-103b was detected by \cite{cartier17}

	spin-orbit alignment: 2016ApJ...823...29A

Thermal emission spectra of other super-hot Jupiters

The atmosphere of WASP-103b has been characterized with transmission and emission spectroscopy \citep{southworth16, cartier17, lendl17}. The thermal emission spectrum is featureless

%, a hot Jupiter with a mass of X, radius of Y, period of Z (cite discovery paper). These properties are similar to WASP-43b, except that WASP-103b has a much (FIXME vs FIXME K) thanks to its hot F?? star host. By comparing the phase curves of these two planets, we can test the effect of irradiation on the climate of the planet.

@@

\subsection{\HST/WFC3}
The \HST\ phase curve observations consisted of two visits on 26-27 February and 2-3 August 2015. Each visit was 15 orbits in duration and spanned 23 hours. The last half of orbit 15 in each visit was used for a gyro bias update and does not produce useable science data.  We took a direct image of the star with the F126N filter at the beginning of each orbit to determine the wavelength solution zero-point. The remainder of the orbit consisted of time-series spectroscopy with the G141 grism ($1.1 - 1.7$ $\mu$m) and the 256 x 256 pixel subarray. We used the SPARS10/NSAMP = 15 read-out mode, which has an exposure time of 103 seconds. To optimize the duty cycle of the observations, we used the spatial scan observing mode with a scan rate of 0.03 arcsec/s, alternating between forward and backward scanning on the detector. The scan height was 25 pixels and the peak counts were $35\times10^3$ photoelectrons per pixel. We collected a total of 18 spatial scan exposures per orbit.  The two eclipse observations from Program 13660 had a similar observing setup.

We reduced the data from both programs using a custom pipeline developed for past analyses of WFC3 data \citep[for details see][]{kreidberg14a, kreidberg14b, kreidberg15b}. Briefly, we use the optimal extraction algorithm of \cite{horne86} to extract each up-the-ramp sample (or ``stripe") separately. The stripes are then summed to create the final spectrum. For each stripe, the extraction window is 24 pixels high and centered on the stripe midpoint. We estimate the background from the median of a region of the detector that is uncontaminated by the target spectrum (rows 5-50). The typical background counts are low (10-15 photoelectrons per pixel, roughly 0.03\% of the peak counts from the target star). We note that the extracted spectrum includes flux from a nearby star, which is separated from WASP-103 by less than two pixels \citep[0.2";][]{wollert15}. Our extracted spectrum includes flux from this star, which we account for later in the analysis.

\subsection{\Spitzer}
We also obtained \Spitzer/IRAC observations with 3.6 and 4.5 $\mu$m photometric filters (referred to as Channel 1 and Channel 2, respectively). The observations had the following setup. Each phase curve observation consisted of 30 hours of time series photometry, beginning three hours prior to one secondary eclipse and ending three hours after a second eclipse.  We used 12 s exposures to maximize the duty cycle without saturating the detector. The data volume is relatively low for this exposure time, so we read out the full array. To minimize the intrapixel effect (variations in flux caused by imprecise pointing), we did not dither and also used PCRS peak-up to improve the pointing accuracy. We began each observations with a 30-minute position settling period, followed by three Astronomical Observation Requests (AORs) of equal duration. At the beginning of each AOR, the telescope was repointed to position the target in the ``sweet spot" of the detector.

@@

The photometric analyses are summarized for each observing season in Table\,\ref{tab:photometry}.  The standard deviation of WASP-103's brightness in each season is given in column~4.  The mean of the three standard deviations is 0.0058~mag, which is very comparable to the mean standard deviation of the six comparison stars (0.0055~mag).  To maximize the possibility of detection WASP-103's rotation, we normalized the photometry such that each observing season has the same mean, thereby removing long-term variability in WASP-103 and/or the comparison stars. We performed a periodogram analysis of the normalized dataset. Figure\,\ref{fig:photometry} shows the frequency spectrum and the phase curve computed with the best frequency, respectively.  We also performed a least-squares fit to the data to determine the best fit sine curve. The best fit period is 6.814 days, which agrees closely with the estimated stellar rotation period of 6.855 days \citep{getal2014}.  The variability amplitude is 0.005~mag, which has FIXME impact on our analysis here.

@@

We calculated the transit model $T(t)$ with the \texttt{batman} package \citep{kreidberg15a}. Many of the physical parameters are tightly constrained by \cite{southworth15}, so we fixed the orbital period, time of inferior conjunction, orbital inclination, and ratio of semi-major axis to stellar radius to the previously published values ($P = 0.925545613$ day, $t_0 = 2456836.2964455\,\mathrm{BJD_{TDB}}$, $i = 87.3^\circ$, and $a/R_s = 2.999$). As a test, we fit for these parameters with the Spitzer Channel 2 light curve, which has the best phase coverage and least systematic noise of the three data sets. We found that the transit parameters are consistent with the \cite{southworth15} results, so we proceeded with the remainder of the analysis holding those parameters fixed.  The free parameters for the transit model were the transit depth $r_p$ and a linear limb darkening parameter $u$.

@@

We scaled the planet-to-star flux by a correction factor $c$ to account for dilution from the companion star and ellipsoidal variability due to the planet's tidal distortion. The correction factor took the form: 
\begin{equation}
	c(\lambda, t) = [1 + \alpha(\lambda)]A(t)
\end{equation}
where $\alpha(\lambda)$ is the additional fractional flux from the companion star and $A(t)$ is the sky-projected area of the planet. We estimated $\alpha(\lambda)$ based on the best fit spectral energy distribution from \cite{cartier17}. The companion star contributes $10-16\%$ of the flux over the wavelength range $1 - 5\,\mu$m.  We calculated $A(t)$ using the analytic formula from \cite{leconte11b}, equation B.9, which computes the projected area of a triaxial ellipsoid. We estimated the ellipsoid properties using Table B.3 of \cite{leconte11a}, assuming the planet radius is 1.5 $R_\mathrm{Jup}$ and age is 5 Gyr. The predicted ellipsoidal variability is shown in Figure\,\ref{fig:ellipsoidal}. At quadrature, the projected radius is $8\%$ larger than at phase zero (mid-transit). %FIXME: say this doesn't matter.
%In addition to temperature, the planet's visible surface area $[R(t)]^2$ is also expected to vary with orbital phase due to tidal distortion \citep{gillon14}. In the observed light curve, changes in planet radius are degenerate with changes in temperature.

@@

\section{Results}
The fitted light curves are shown in Figure\,\ref{fig:phasecurves}. This figure shows results from the kinematic model for the thermal phase variation and has instrument systematics removed.  Broadly speaking, the phase curves show large amplitude phase variation, ranging from 1.5 parts-per-thousand (ppt) in the WFC3 bandpass to 5 ppt in \Spitzer\ Channel 2.  The planet flux changes significantly with orbital phase in all three of the data sets, suggesting a strong gradient from dayside to nightside temperature. The peak brightness occurs near phase 0.5.
In this section we discuss the quality

@@

@@

\subsection{Blackbody Fits}
A blackbody is the simplest model for the planet's thermal emission and provide a useful first evaluation of the atmospheric properties. The best fit blackbodies are shown in Figures \ref{fig:spectra} and \ref{fig:dayside}. This model performs well for some orbital phases, with $\chi^2_\nu$ ranging from 0.9 to 1.1 for phases $\phi = 0.1, 0.7$ and $0.8$. At other phases the blackbody model is a poorer fit, particularly to the \Spitzer\ data. For phases $0.2 - 0.6$, the \Spitzer\ eclipse depths are larger than predicted for the best fit blackbody, suggesting the presence of an emission feature. On the nightside (phase $0.9$), the \Spitzer\ $4.5\,\mu$m eclipse depth is \emph{smaller} than predicted, suggesting an absorption feature. These differences hint at changing thermal structure with longitude in the atmosphere.

@@

There are many factors contributing to the blackbody spectra --
weak water xsec contrast, isothermal pt, H-, and dissociated water

@@

>>>>

\begin{abstract}
	We present thermal phase curve measurements for the hot Jupiter WASP-103b observed with \emph{Hubble}/Wide Field Camera 3 and \emph{Spitzer}/IRAC. The dayside planet-to-star flux ranges from FIXME to FIXME at $1.1 - 4.5\,\mu$m.  The phase curves are symmetric and large amplitude, indicative of poor heat redistribution. We fit the phase curves with the SPIDERMAN package to determine the brightness distribution in the photosphere, and find the peak dayside temperature ranges from FIXME to FIXME depending on model.  We also generate phase-resolved spectra, which are consistent with blackbody models for the planet's emission at near-infrared wavelengths, but show strong emission features in the \Spitzer\ bands. We attribute these features to CO emission and a thermal inversion. The retrieved atmospheric composition is more metal-enriched than other hot Jupiters (FIXME). The carbon-to-oxygen ratio is consistent with solar composition. The absence of water features is due to BLAH. We compare the phase curves to general circulation models and find that atmospheric drag is required.  We also compare the spectra to brown dwarfs and young directly imaged companions and find that the water features are significantly diminished in amplitude. power of phase curve observations to elucidate the three-dimensional structure of exoplanet atmospheres.

inversion on the dayside but not on the nightside
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
Planets are round, rotating, and irradiated on one hemisphere at a time -- all of which contribute to rich spatial structure in their climate and atmospheric composition.  Short period, tidally locked planets are an extreme example, with hot, continuously illuminated daysides. This asymmetry is expected to produce large gradients in temperature, chemistry, and cloud coverage with longitude \citep{showman09, kataria16, parmentier16}, and provides an opportunity to learn about atmospheric dynamics in a very different regime from the Solar System planets.

Exoplanets are so distant that they are not spatially resolved, so the only way to reveal inhomogeneity in their atmospheres is to measure a phase curve. Phase curve observations consist of continuous monitoring of the planet-to-star flux ratio over a complete orbital revolution of the planet, which exposes different longitudes at each orbital phase.  The first phase curve was observed with \Spitzer\ for the hot Jupiter HD 189733b by \citep{knutson07}, followed by additional \Spitzer\ observations for about a dozen more systems \citep[cataloged in][]{parmentier17}.  The observations revealed large day-night temperature contrasts (in excess of 300 K), and eastward shifted peak brightness due to heat circulation, as predicted by 3D models \citep{showman09}.  These infrared measurements were complemented by optical phase curves from \Kepler\, which show evidence for reflected light from patchy and variable dayside clouds with a range of compositions \citep{demory13, hu15, armstrong16, parmentier16}. A \emph{spectroscopic} phase curve was observed for WASP-43b with \emph{Hubble}/Wide Field Camera 3 (\HST/WFC3) in the near-infrared, which provided the first phase-resolved measurements of an exoplanet's water abundance and thermal structure \citep{stevenson14, stevenson17}.

In this paper we present spectroscopic phase curve observations of the hot Jupiter WASP-103b, measured with \HST/WFC3 and Spitzer/IRAC. This planet is an ideal target for phase curve observations, with an orbital period of just 22 hours and an equilbrium temperature of 2500 K \citep{gillon14}.  Previous observations of WASP-103b's atmosphere revealed a blackbody-like dayside emission spectrum, with possible evidence for a $K_\mathrm{S}$-band emission feature \citep{cartier17, delrez18}. The shape of the optical transmission spectrum is disputed, with some evidence for strong sodium and potassium absorption features that are consistent with expectations for a cloud-free atmosphere \citep{lendl17}, but other observations indicating a large optical scattering slope that is not fit well by theoretical models \citep{southworth16}.

WASP-103b is an archetype of the class of ultra-hot Jupiters (with dayside temperatures $>2000\,\mathrm{K}$). Observations of the near-IR thermal emission spectra of these objects have been puzzling: some show absorption features, some show emission features, and some are consistent with blackbody models \citep{madhusudhan11, crossfield12, stevenson14b, haynes15, evans16, beatty17a, beatty17b, }.  A variety of explanations have been proposed, including low metallicity or carbon-to-oxygen ratios \citep{madhusudhan11, cartier17}, which would TiO/VO heating of the upper atmosphere, .

Our phase-curve observations of provide the first 3D constraints on temperature structure of a planets.

%	spin-orbit alignment: 2016ApJ...823...29A

%are featureless (consistent with a blackbody model for the planet), others show water absorption features, and water emission \citep{evans16}. Isothermal in the near-IR \citep{crossfield12, stevenson14, }. Some have absortion features in near-IR \citep{kreidberg14b, line16}

Fortney 08

Phase curve observation to provide context for the global thermal structure of the planet.

@@

\subsection{\HST/WFC3}
The \HST\ phase curve observations consisted of two visits on 26-27 February and 2-3 August 2015. Each visit was 15 orbits in duration and spanned 23 hours. The last half of orbit 15 in each visit was used for a gyro bias update and does not produce useable science data.  We took a direct image of the star with the F126N filter at the beginning of each orbit to determine the wavelength solution zero-point. The remainder of the orbit consisted of time-series spectroscopy with the G141 grism ($1.1 - 1.7$ $\mu$m) and the 256 x 256 pixel subarray. We used the SPARS10/NSAMP = 15 read-out mode, which has an exposure time of 103 seconds. To optimize the duty cycle of the observations, we used the spatial scan observing mode with a scan rate of 0.03 arcsec/s, alternating between forward and backward scanning on the detector. The scan height was 25 pixels and the peak counts were $35\times10^3$ photoelectrons per pixel. We collected a total of 18 spatial scan exposures per orbit.  The two eclipse observations from Program 13660 had a similar observing setup \citep[described in detail in][]{cartier17}.

We reduced the data from both programs using a custom pipeline developed for past analyses of WFC3 data \citep[for details see][]{kreidberg14a, kreidberg14b, kreidberg15b}. Briefly, we use the optimal extraction algorithm of \cite{horne86} to extract each up-the-ramp sample (or ``stripe") separately. The stripes are then summed to create the final spectrum. For each stripe, the extraction window is 24 pixels high and centered on the stripe midpoint. We estimate the background from the median of a region of the detector that is uncontaminated by the target spectrum (rows 5-50). The typical background counts are low (10-15 photoelectrons per pixel, roughly 0.03\% of the peak counts from the target star). We note that the extracted spectrum includes flux from a nearby companion star, which is separated from WASP-103 by less than two pixels \citep[0.2";][]{wollert15}. Our extracted spectrum includes flux from this star, which we account for later in the analysis.

\subsection{\Spitzer}
We also obtained \Spitzer/IRAC observations with 3.6 and 4.5 $\mu$m photometric filters (referred to as Channel 1 and Channel 2, respectively). The observations had the following setup. Each phase curve observation consisted of 30 hours of time series photometry, beginning three hours prior to one secondary eclipse and ending three hours after a second eclipse.  We used 12 s exposures to maximize the duty cycle without saturating the detector. The data volume is relatively low for this exposure time, so we read out the full array. To minimize the intrapixel effect (variations in flux caused by imprecise pointing), we did not dither and also used PCRS peak-up\footnote{\url{http://irsa.ipac.caltech.edu/data/SPITZER/docs/irac/pcrs\_obs.shtml}} to improve the pointing accuracy. We began each observations with a 30-minute position settling period, followed by three Astronomical Observation Requests (AORs) of equal duration. At the beginning of each AOR, the telescope was repointed to position the target in the ``sweet spot" of the detector.

@@

The photometric analyses are summarized for each observing season in Table\,\ref{tab:photometry}.  The standard deviation of WASP-103's brightness in each season is given in column~4.  The mean of the three standard deviations is 0.0058~mag, which is very comparable to the mean standard deviation of the six comparison stars (0.0055~mag).  To maximize the possibility of detection WASP-103's rotation, we normalized the photometry such that each observing season has the same mean, thereby removing long-term variability in WASP-103 and/or the comparison stars. We performed a periodogram analysis of the normalized dataset. Figure\,\ref{fig:photometry} shows the frequency spectrum and the phase curve computed with the best frequency, respectively.  We also performed a least-squares fit to the data to determine the best fit sine curve. The best fit period is 6.814 days, which agrees closely with the estimated stellar rotation period of 6.855 days \citep{gillon14}.  The variability amplitude is 0.005~mag, which has FIXME impact on our analysis here.

@@

We calculated the transit model $T(t)$ with the \texttt{batman} package \citep{kreidberg15a}. Many of the physical parameters are tightly constrained by \cite{southworth15}, so we fixed the orbital period, time of inferior conjunction, orbital inclination, and ratio of semi-major axis to stellar radius to the previously published values ($P = 0.925545613$ day, $t_0 = 2456836.2964455\,\mathrm{BJD_{TDB}}$, $i = 87.3^\circ$, and $a/R_s = 2.999$). As a test, we fit for these parameters with the \Spitzer\ Channel 2 light curve, which has the best phase coverage and least systematic noise of the three data sets. We found that the transit parameters are consistent with the \cite{southworth15} results, so we proceeded with the remainder of the analysis holding those parameters fixed.  The free parameters for the transit model were the transit depth $r_p$ and a linear limb darkening parameter $u$.

@@

We scaled the planet-to-star flux by a correction factor $c$ to account for dilution from the companion star and ellipsoidal variability due to the planet's tidal distortion. The correction factor took the form: 
\begin{equation}
	c(\lambda, t) = [1 + \alpha(\lambda)]A(t)
\end{equation}
where $\alpha(\lambda)$ is the additional fractional flux from the companion star and $A(t)$ is the sky-projected area of the planet. We estimated $\alpha(\lambda)$ based on the best fit spectral energy distribution from \cite{cartier17}. The companion star contribution ranges from $10\%$ at $1.1\,\mu$m to $16\%$ at $4.5\,\mu$m. (FIXME: what is the error?)  We calculated $A(t)$ using the analytic formula from \cite{leconte11b}, equation B.9, which computes the projected area of a triaxial ellipsoid. We estimated the ellipsoid properties using Table B.3 of \cite{leconte11a}, assuming the planet radius is 1.5 $R_\mathrm{Jup}$ and age is 5 Gyr. The predicted ellipsoidal variability is shown in Figure\,\ref{fig:ellipsoidal}. At quadrature, the projected radius is $8\%$ larger than at phase zero (mid-transit). %FIXME: say this doesn't matter.
%In addition to temperature, the planet's visible surface area $[R(t)]^2$ is also expected to vary with orbital phase due to tidal distortion \citep{gillon14}. In the observed light curve, changes in planet radius are degenerate with changes in temperature.

@@

\section{Results}
The fitted light curves are shown in Figure\,\ref{fig:phasecurves}. This figure shows results from the spherical harmonics model for the thermal phase variation and has instrument systematics removed.  Broadly speaking, the phase curves show large amplitude phase variation, ranging from 1.5 parts-per-thousand (ppt) in the WFC3 bandpass to 5 ppt in \Spitzer\ Channel 2.  The planet flux changes significantly with orbital phase in all three of the data sets, suggesting a strong gradient from dayside to nightside temperature, and peak brightness occurs near phase 0.5. In this section, we describe the phase curve fits and the phase-resolved spectra. FIXME

@@

\subsection{Comparison with Previous Observations}
Cartier, Delrez.

@@

\subsection{Blackbody Fits}
A blackbody is the simplest model for the planet's thermal emission and provide a useful first evaluation of the atmospheric properties. The best fit blackbodies are shown in Figures \ref{fig:spectra} and \ref{fig:dayside}. This model performs well for some orbital phases, with $\chi^2_\nu$ ranging from 0.9 to 1.1 for phases $\phi = 0.1, 0.7$ and $0.8$. At other phases the blackbody model is a poorer fit, particularly to the \Spitzer\ data. For phases $0.2 - 0.6$, the \Spitzer\ eclipse depths are larger than predicted for the best fit blackbody, suggesting the presence of an emission feature. On the nightside (phase $0.9$), the \Spitzer\ $4.5\,\mu$m eclipse depth is \emph{smaller} than predicted, suggesting an absorption feature. These differences suggest changing thermal structure with longitude in the atmosphere.

@@

There are many factors contributing to the blackbody spectra --
weak water xsec contrast, isothermal pt, H-, and dissociated water, gravity (cite T. Beatty)

@@

eclipse mapping to break degeneracy in models

We rule out the low metallicity and or high C/O models

confirm prediction that TiO is present in the hottest planets
